
\subsection{}

\begin{exercise}
  What is a convex set?
  What is an affine set?
  Examples:\@ the unit cube, hyperplanes.
\end{exercise}

\begin{solution}
  A convex set is intuitively a set where for each pair of two
  points, all the points one the line between those two points are
  also in the set.
  This for instance means that the unit circle is convex, while a
  typical drawing of a star is not, as connecting two ``points''
  places you outside the star.
  Explicitly, we have that a set $C$ is convex if and only if
  \begin{equation}
    \lambda x + (1 - \lambda) y \in C
  \end{equation}
  for all $x, y \in C$ and $\lambda \in [0, 1]$.

  An affine set is practically a linear subspace centred around some
  point other than zero.
  It is an affine set if and only if it is the solution set to some
  linear equation.
  That is, for a non-empty set $S \subseteq \R^n$, there is a matrix
  $A  \in \R^{m,n}$ and a vector $b \in \R^m$ such that
  \begin{equation}
    S = \{ x \in \R^n : Ax = b \}.
  \end{equation}
  It can equivalently be written as $C = L + x_0 = \{ x + x_0 : x \in
  L \}$ for some linear subspace $L$ of $\R^n$.

  To see the equivalence, let $x_0$ be a solution of the linear
  equation, such that $Ax_0 = b$.
  Then, $L$ is the kernel of $A$, as then for $z \in S$ we have
  \begin{equation}
    Az = A(x + x_0) = Ax + Ax_0 = b \implies Ax = 0.
  \end{equation}
\end{solution}

\begin{exercise}
  Is the sum of convex sets also convex?
\end{exercise}

\begin{solution}
  Yes.
  Consider two convex sets $A,B \subseteq \R^n$.
  We then have that the sum of the sets is given by
  \begin{equation}
    A + B = \{ a + b : a \in A, b \in B \}.
  \end{equation}
  As $A$ and $B$ are convex, we have for $a_1, a_2 \in A$, $b_1, b_2
  \in B$ and $\lambda \in [0, 1]$ that
  \begin{equation}
    \lambda a_1 + (1 - \lambda) a_2 + \lambda b_2 + (1 - \lambda) b_2
    = \lambda (a_1 + b_1) + (1 - \lambda) (a_2 + b_2)
    \in A + B.
  \end{equation}
\end{solution}

\begin{exercise}
  What is a cone?
  Example:\@ First octant in $\R^3$.
\end{exercise}

\begin{solution}
  A cone $C$ is a set with the property that a non-negative
  combination of two points in the cone, is also in the cone.
  That is, for $\lambda_1, \lambda_2 \geq 0$ and $x, y \in C$, we have
  \begin{equation}
    \lambda_1 x + \lambda_2 y \in C.
  \end{equation}
  The first octant in $\R^3$, often denoted $\R^3_+$, is an example
  of such a space.
\end{solution}

\begin{exercise}
  What is a polyhedron?
\end{exercise}

\begin{solution}
  A polyhedron is the solution set to any linear equation $Ax \leq b$.
\end{solution}
